### 2017-06-28 doc/RMD/RDA-analysis.Rmd
changed all svg figures to pdf

---
title: "dbRDA analysis"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

This document serves to re-assess Sajeewa's dbRDA analysis with the repeat
lengths that have been modified to avoid inconsistent allele calls.

Modification in this case means that a tiny amount has been added or subtracted
to the repeat length to ensure that the alleles are all unique after division.

Below is my attempt at reproduction of Sajeewa's analysis.

## Packages and Data

```{r, load_packages}
library('tidyverse')
library('poppr')
library('vegan')
library('ggrepel')
```

So, here Sajeewa clone-corrected the data according to the combination of
Host, Source (aka Field), Region (aka State/Country), and Year.

```{r, load_data}
load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
setPop(dat)   <- ~Host/Source/Region/Year
setPop(dat11) <- ~Host/Source/Region/Year
dat11cc <- clonecorrect(dat11, ~Host/Source/Region/Year, keep = 1:4)
dat16cc <- clonecorrect(dat, ~Host/Source/Region/Year, keep = 1:4)
dat11cc
dat16cc
# Asserting that nothing messed up with the metadata.
stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
stopifnot(identical(indNames(dat16cc), other(dat16cc)$meta$Isolate))

# function to gather Environmental variables, but averaging Severity
# (aka Straw Test, Virulence). This function is necessary because the
# data tends to randomly shuffle when being processed vai dplyr functions
# for better or for worse. ¯\_(ツ)_/¯
# 
# @param DAT the full data set
# @param CC The clone-corrected data set
# @param wmn should the "Source" column be converted to a binary wmn
#   factor? When this is true, everything that is not wmn will be converted
#   to "other"
# 
# @return A data frame containing Severity, Host, Source, Year, Region, and MLG
makeENV <- function(DAT, CC, wmn = FALSE){
  # Creating the data frame with severity
  META   <- select(other(DAT)$meta, -Isolate)
  STRATA <- select(strata(DAT), -MCG)
  STRATA <- if (wmn) mutate(STRATA, Source = ifelse(Source == "wmn", "wmn", "other")) else STRATA
  MLL    <- data.frame(MLG = mll(DAT))
  sev    <- bind_cols(META, STRATA, MLL) %>%
    group_by(Host, Source, Year, Region, MLG) %>% 
    summarize(Severity = mean(Severity)) %>% # Get mean severity per MLG
    ungroup()
  # Ensuring the data is in the correct order
  META   <- select(other(CC)$meta, -Isolate)
  STRATA <- select(strata(CC), -MCG)
  MLL    <- data.frame(MLG = mll(CC))
  bind_cols(META, STRATA, MLL) %>% 
    left_join(sev)
}

# Get environmental variables for 11 loci
ENV11 <- makeENV(dat11, dat11cc)
ENV11
stopifnot(identical(ENV11$MLG, mll(dat11cc)))
ENV11 <- select(ENV11, -MLG)

# Get environmental variables for 16 loci
ENV16 <- makeENV(dat, dat16cc)
ENV16
stopifnot(identical(ENV16$MLG, mll(dat16cc)))
ENV16 <- select(ENV16, -MLG)
```

## Functions to tie everything together

The analysis has a couple of steps

1. Model choice. Since we don't want to overparamaterize the model, we will use 
*vegan*'s built in model choice function `ordistep()` to to forward-backward 
selection of the appropriate model to fit our data. Yes, some (particularly 
Bayesians) believe that model choice is evil and the proper analysis will find 
the true pattern in the underlying data, but hey, I'm just trying to make sure 
that I'm not making overconfident judgements.

2. Plot the results of the model choice using ggplot2, overlaying the top 8 
explanitory vectors

```{r modelfun}
# model choice for dbrda/capscale
# 
# @param bdist a distance matrix (here, we use Bruvo's distances, which is why
#        it's called 'bdist').
# @param ENV a data frame of environmental variables with the same number of
#        observatios as bdist
# @param CHOOSER the name of the function to perform the forwards-backwards selection
# @param ... arguments to be passed on to CHOOSER
# 
# @return a capscale object
choose_dbrda <- function(bdist, ENV, CHOOSER = "ordistep", ...){
  # Step 1: create null model
  mod0  <- capscale(bdist ~ 1, data = ENV, add = TRUE)
  # Step 2: create full model (all variables in ENV)
  mod1  <- capscale(bdist ~ ., data = ENV, add = TRUE)
  # Step 3: Run forward-backwards selection (this can take a while)
  CHOOSER   <- match.fun(CHOOSER)
  the_model <- CHOOSER(mod0, scope = formula(mod1), ...)
  # Return the best model with the anova results
  return(the_model)
}

# This was ganked from https://github.com/gavinsimpson/ggvegan/blob/59d233977a5b2d15d4de150b782fb1794aa1de8b/R/utils.R
# 
# @title Scale Vectors to Data
# @description Scale vector arrows to \code{fill} proportion of the data.
# @param arrows a two-column matrix-like object containing coordinates for the arrows/vectors on x and y axes.
# @param data a two-column matrix-like object containing coordinates of the data on the x and y axes.
# @param at numeric vector of length 2; location of the origin of the arrows.
# @param fill numeric; what proportion of the range of the data to fill
# @return a numeric multiplier that will scale the arrows
# @author Gavin L. Simpson
arrowMul <- function(arrows, data, at = c(0, 0), fill = 0.75) {
    u <- c(range(data[,1], range(data[,2])))
    u <- u - rep(at, each = 2)
    r <- c(range(arrows[, 1], na.rm = TRUE), range(arrows[, 2], na.rm = TRUE))
    rev <- sign(diff(u))[-2]
    if (rev[1] < 0)
        u[1:2] <- u[2:1]
    if (rev[2] < 0)
        u[3:4] <- u[4:3]
    u <- u/r
    u <- u[is.finite(u) & u > 0]
    fill * min(u)
}
# Plotting the dbRDA results
# 
# @param db a capscale object
# @return a ggplot2 object from the scores 
plot_dbrda <- function(db, arrows = 10, seed = 2017-06-28){
  set.seed(seed)
  dbsum     <- scores(db, display = c("cn", "bp", "sites"), scaling = "sites")
  Centroids <- as.data.frame(dbsum$centroids)
  Centroids <- rownames_to_column(Centroids, var = "cent_type")
  Centroids <- mutate_(Centroids, .dots = list(Length = ~sqrt(CAP1^2 * CAP2^2)))
  # Centroids
  SampleCentroids <- rownames_to_column(data.frame(dbsum$sites), var = "isolate_names")
  labs    <- summary(db, axes = 0)[["cont"]][["importance"]]["Proportion Explained", 1:2]
  terms   <- paste0("(", paste(attr(db$terms, "term.labels"), collapse = "|"), ")")
  mul     <- arrowMul(dbsum$biplot[, 1:2], dbsum$sites)
  Arrows  <- data.frame(dbsum$biplot * mul)
  Arrows  <- rownames_to_column(Arrows, var = "class")
  # Making the classes presentable
  Arrows$class <- gsub(terms, "\\1: ", Arrows$class)
  Arrows$class <- gsub(": $", "", Arrows$class)
  Arrows$class <- gsub("unk", "unknown", Arrows$class)
  Arrows  <- mutate_(Arrows, .dots = list(Length = ~sqrt(CAP1^2 * CAP2^2)))
  Arrows  <- arrange(Arrows, Length)
  Arrows  <- top_n(Arrows, arrows)
  ggplot(Centroids, aes(x = CAP1, y = CAP2))+
    geom_point(data = SampleCentroids, 
               # alpha = 1/2, 
               # fill = "white",
               # fill = "dark orange", 
               color = "grey45", 
               size = 2.5, 
               pch = 21)+
    coord_cartesian() +
    geom_segment(aes(x = 0, xend = CAP1, 
                     y = 0, yend = CAP2),
                 arrow = arrow(length = unit(0.3, "cm")), 
                 data = Arrows
                 ) + 
    geom_label_repel(aes(x = CAP1, y = CAP2, label = class), 
                     point.padding = unit(0.5, "lines"),
                     segment.color = "grey25",
                     data = Arrows) +
    xlab(paste0("Eig 1 (", round(labs[[1]]*100, 2), "% variance explained)")) +
    ylab(paste0("Eig 2 (", round(labs[[2]]*100, 2), "% variance explained)"))
}
```

# Calculations

Here come the calculations. Note, since I am verifying that we get the same
results from the 16 loci as we do for the 11 loci, we have to do this twice.

1. calculate Bruvo's genetic distance.
2. model choice

```{r rda_calculation, cache = TRUE}
# 11 loci
dat11cc.bruvo <- dat11cc %>% bruvo.dist(replen = other(.)$REPLEN)
cap11cc       <- choose_dbrda(dat11cc.bruvo, ENV = ENV11, CHOOSER = "ordistep")
# 16 loci
dat16cc.bruvo <- dat16cc %>% bruvo.dist(replen = other(.)$REPLEN)
cap16cc       <- choose_dbrda(dat16cc.bruvo, ENV = ENV16, CHOOSER = "ordistep")
```

# ANOVA

The results of our modeling and the marginal effects for each parameter on the
model itself.

```{r}
cap11cc
cap11cc$anova
cap16cc
cap16cc$anova
```

We can test the effects of how much of the model explains the variance

```{r total_anova, cache = TRUE}
anova(cap11cc)
anova(cap16cc)
```

We can also see the marginal effects with `by = "margin"`

```{r marginal_anova, cache = TRUE}
anova(cap11cc, by = "margin")
anova(cap16cc, by = "margin")
```

# Plot the results

```{r resultplot, fig.width = 5, fig.height = 5}
set.seed(99)
plot_dbrda(cap11cc) + 
  theme_classic(base_size = 16, base_family = "Helvetica") + 
  theme(axis.text = element_text(color = "black")) +
  theme(aspect.ratio = 1)
FILE <- file.path(PROJHOME, "results", "figures", "publication", "Figure7Z.pdf")
ggsave(filename = FILE, width = 88, height = 88, units = "mm", scale = 1.5)
plot_dbrda(cap16cc) +
  theme_classic(base_size = 16, base_family = "Helvetica") + 
  theme(axis.text = element_text(color = "black")) +
  theme(aspect.ratio = 1)
```

This is all well and good, but what exactly does it actually mean to have only 
1.93 variance explained (for 11 loci, that is)? I was a bit stumped as well. 
Luckily, Gavin Simpson has some answers: http://stackoverflow.com/a/22537820/2752888

Basically, we have a situation where we have all of the variables only
explaining a whopping 16 - 17% of the variance and the rest goes unexplained.

```{r, echo = FALSE, results = "asis"}
cap11sum <- summary(cap11cc, axes = 0, scaling = "sites")
cap16sum <- summary(cap16cc, axes = 0, scaling = "sites")
vartab <- function(x){
  tot <- x$tot.chi
  con <- x$constr.chi
  uco <- x$unconst.chi
  data.frame(list(Inertia = c(tot, con, uco),
                  Proportion = c(tot, con, uco)/tot))
}
cat("\n\n11 loci:\n\n")
knitr::kable(vartab(cap11sum), digits = 3)
cat("\n\n16 loci:\n\n")
knitr::kable(vartab(cap16sum), digits = 3)
```

## Variance partitioning

First, we want to examine how the entire data fits with the model. We will use
the alleles as the representative of the entire data.

```{r}
dat11raw <- genind2df(dat11cc, usepop = FALSE) %>% mutate_all(funs(as.integer(.)))
dat16raw <- genind2df(dat16cc, usepop = FALSE) %>% mutate_all(funs(as.integer(.)))
vp11     <- varpart(dat11cc.bruvo, ~Severity + Year + Region + Host, data = ENV11, comm = dat11raw, add = TRUE)
plot(vp11, Xnames = c("Full Model", "No Model"))
try(vp16     <- varpart(dat16cc.bruvo, ~Year + Region + Host, data = ENV16, comm = dat16raw, add = TRUE))
```

```{r}
vp11 <- varpart(dat11cc.bruvo, ~Severity, ~Year, ~Region, ~Host, data = ENV11, add = TRUE)
vp16 <- varpart(dat16cc.bruvo, ~Year, ~Region, ~Host, data = ENV16, add = TRUE)

plot(vp11, digits = 2, Xnames = c("Severity", "Year", "Region", "Host"))

plot(vp16, digits = 2, Xnames = c("Year", "Region", "Host"))
```

Out of both of the models, only Region explains the most variance, almost double
that of Host or Year and Severity contributes almost nothing.

```{r}
wmn11 <- seppop(dat11, ~Source)$wmn %>% setPop(~Region)
wmn11_sub <- popsub(wmn11, blacklist = character(0))
min_span_net <- bruvo.msn(wmn11_sub, replen = other(wmn11)$REPLEN, add = TRUE, loss = TRUE, showplot = FALSE, include.ties = TRUE)
set.seed(67)
plot_poppr_msn(wmn11,
               min_span_net,
               inds = "none",
               mlg = TRUE,
               gadj = 6,
               nodelab = 100,
               nodebase = 1.15,
               palette = funky,
               cutoff = (0.5/11)*3,
               quantiles = FALSE,
               beforecut = TRUE,
               layfun = igraph::layout_with_lgl)
```

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>