### 2017-06-30 doc/RMD/wmn-differentiation.Rmd
Add wmn-differentiation

---
title: "R Notebook"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

```{r packages, message = FALSE, warning = FALSE}
library("tidyverse")
library("poppr")
(load("data/sclerotinia_16_loci.rda"))
```

# Purpose

The white mold nursery populations are unique because they are not fungicide
treated and have the same cultivars planted in them year after year.

The question becomes, are white mold nurseries differentiated from each other or
are they more or less homogeneous? We could use AMOVA to test for these with
location and binary source (wmn or non-wmn) as the hierarchy.

# Data Setup

First, we want to clone-correct our data down to the field level so that we
don't accidentally include non-independent samples.

```{r correction}
dat11
dat11cc <- clonecorrect(dat11, ~Region/Source/Host/Year)
dat11cc
```

Now that we've done that, we should make a new variable in the strata that 
separates the white mold nurseries from the others. We'll call this stratum 
"Source Type".

```{r new_stratum}
addStrata(dat11cc) <- strata(dat11cc) %>% 
  mutate(SourceType = forcats::fct_inorder(ifelse(Source == "wmn", "wmn", "other"))) %>%
  select(SourceType)
setPop(dat11cc) <- ~SourceType
dat11cc
```

I can perform AMOVA on the newly defined variable using Bruvo's distance.

```{r AMOVA}
other(dat11cc)$REPLEN
bd <- bruvo.dist(dat11cc, replen = other(dat11cc)$REPLEN)
(ssc_amova <- poppr.amova(dat11cc, ~SourceType, dist = bd, quiet = TRUE))
ssc_amova_test <- randtest(ssc_amova, nrepet = 999)
plot(ssc_amova_test)
ssc_amova_test
```

This result is telling us that there is some subdivision between white mold
nurseries and non-white mold nurseries. Of course, from previous analyses, we
know that Mexico is differentiated from other populations, so what happens if we
account for Region? Here, we are placing region lower in the heirarchy because
we specifically want to test the effect of region on the differentiation between
white mold nurseries within different regions.

```{r AMOVA-Region, cache = TRUE}
(ssc_amova_region <- poppr.amova(dat11cc, ~SourceType/Region, dist = bd, quiet = TRUE))
ssc_amova_region_test <- randtest(ssc_amova_region, nrepet = 999)
plot(ssc_amova_region_test)
ssc_amova_region_test
```

Okay! This shows that when we account for Region after accounting for Source
Type, we find that the differentiation is coming mainly from the Regions. What
happens when we remove Mexico?

```{r AMOVA-nomex, cache = TRUE}
datnomex <- setPop(dat11cc, ~Region) %>% popsub(blacklist = "Mexico")
bdnm     <- bruvo.dist(datnomex, replen = other(datnomex)$REPLEN)
(ssc_amova_nm <- poppr.amova(datnomex, ~SourceType/Region, dist = bdnm, quiet = TRUE))
ssc_amova_nm_test <- randtest(ssc_amova_nm, nrepet = 999)
plot(ssc_amova_nm_test)
ssc_amova_nm_test
```

When we remove the Mexican isolates (which only contained white mold nurseries
and shared no genotypes), we see that indeed, the degree of differentiation
went down.

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>