### 2017-07-06 doc/RMD/pop-diff.Rmd
made compound dapc figure

---
title: "Population Differentiation Analysis"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
knitr::opts_chunk$set(fig.align = "center")
```

## Packages and Data

```{r load_packages, warning = FALSE, message = FALSE}
library('tidyverse')
library('poppr')
library('ggcompoplot')
```

```{r load_data}
load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
setPop(dat11) <- ~Host/Source/Region/Year
dat11cc <- clonecorrect(dat11, ~Host/Source/Region/Year, keep = 1:4)
dat11cc

# Asserting that nothing messed up with the metadata.
stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
dat11.bruvo   <- bruvo.dist(dat11, replen = other(dat11)$REPLEN)
dat11cc.bruvo <- bruvo.dist(dat11cc, replen = other(dat11)$REPLEN)
```

# Genotypic Diversity

```{r geno-div, cache = TRUE, fig.width = 10, fig.height = 10, comment = NA}
ptab <- dat11cc %>% setPop(~Region) %>% poppr(sample = 999, total = FALSE)
```

```{r geno-div2, fig.width = 10, fig.height = 10, comment = NA}
ptab %>%
  arrange(desc(N), desc(MLG)) %>%
  mutate(Ia = case_when(
    !is.finite(.$p.Ia) ~ "-",
    .$p.Ia == 0.001 ~ paste0(round(.$Ia, 2), "*"),
    .$p.Ia <= 0.01 ~  paste0(round(.$Ia, 2), "-"),
    .$p.Ia <= 0.05 ~  paste0(round(.$Ia, 2), "~"),
    TRUE ~ as.character(round(.$Ia, 2))
  )) %>%
  mutate(rbarD = case_when(
    !is.finite(.$p.rD) ~ "-",
    .$p.rD == 0.001 ~ paste0(round(.$rbarD, 2), "*"),
    .$p.rD <= 0.01 ~  paste0(round(.$rbarD, 2), "-"),
    .$p.rD <= 0.05 ~  paste0(round(.$rbarD, 2), "~"),
    TRUE ~ as.character(round(.$rbarD, 2))
  )) %>%
  mutate_if(is.numeric, round, digits = 2) %>%
  mutate(N = paste(N, MLG, sep = " (") %>% paste0(")") %>% formatC()) %>%
  mutate(eMLG = paste(eMLG, SE, sep = " (") %>%  paste0(")") %>% formatC()) %>%
  rename(`*h*` = Hexp) %>%
  rename(`$E_5$` = E.5) %>%
  rename(`$I_A$` = Ia) %>%
  rename(`$\\lambda$` = lambda) %>%
  rename(`$\\bar{r}_d$` = rbarD) %>%
  select(-File, -MLG, -SE, -p.rD, -p.Ia) %>% 
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(huxtable::everywhere, 4:5, 1) %>%
  huxtable::set_align(huxtable::everywhere, -1, "right") %>%
  huxtable::set_col_width(c(1.1, 0.9, 1.3, 0.5, 0.5, 1.2, 0.7, 0.6, 0.6, 0.8)) %>%
  huxtable::print_md(max_width = 80)

ptab %>% 
  arrange(desc(N), desc(MLG)) %>% 
  write_csv(file.path(PROJHOME, "results/tables/genotype_diversity_table_region.csv"), col_names = TRUE)

mtab <- mlg.table(dat11, ~Region, background = TRUE)
```

# AMOVA

First, we can conduct an AMOVA analysis across all populations, with respect to
year.

```{r amova-no-cc, cache = TRUE}
(res   <- poppr.amova(dat11cc, ~Region/Host/Year, dist = dat11cc.bruvo))
set.seed(999)
(res.p <- randtest(res, nrepet = 999))
```

## DAPC of Region

This can tell us how well our regions separate

```{r dapc-region-calc, fig.width = 8, fig.height = 7, cache = TRUE}
set.seed(2017-07-05)
region.xval <- xvalDapc(tab(dat11cc), strata(dat11cc)$Region, n.pca = 15:25, 
                        n.rep = 100, parallel = "multicore", ncpus = 4)
region.dapc <- region.xval$DAPC
```

```{r dapc-region, fig.width = 9, fig.height = 8, cache = TRUE}
scatter(region.dapc, 
        scree.pca = TRUE,
        bg = "grey95",
        col = "black",
        # col = rev(funky(nlevels(strata(dat11cc)$Region))), 
        pch = seq(nlevels(strata(dat11cc)$Region)),
        legend = "true",
        posi.leg = "topleft",
        posi.pca = "topright"
        )
ggcompoplot(region.dapc, setPop(dat11cc, ~Region), pal = rev(funky(nlevels(strata(dat11cc)$Region))), cols = 2)
```

```{r fig.width = 10, fig.height = 6}
LDS <- bind_cols(Population = region.dapc$grp, as.data.frame(region.dapc$ind.coord)) %>%
  as_tibble()
LDS_pop <- LDS %>% 
  group_by(Population) %>% 
  summarize_all(mean) %>%
  rename_all(function(x) gsub("LD", "mean", x))
LDS <- full_join(LDS, LDS_pop)
LDS_PLOT <- ggplot(LDS, aes(x = LD1, y = LD2, color = Population)) + 
  geom_point(aes(fill = Population), alpha = 0.5, pch = 21, color = "black") +
  geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
  stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
  ggrepel::geom_label_repel(aes(x = mean1, y = mean2, label = Population), 
                            data = LDS_pop, show.legend = FALSE, color = "black") +
  theme_bw() +  
  theme(aspect.ratio = 1/1.618) +
  theme(legend.position = "bottom") +
  theme(axis.text = element_blank()) + 
  theme(axis.title = element_blank()) + 
  theme(axis.ticks = element_blank()) + 
  viridis::scale_color_viridis(discrete = TRUE) +
  viridis::scale_fill_viridis(discrete = TRUE) +
  scale_y_continuous(breaks = 0) + 
  scale_x_continuous(breaks = 0) + 
  theme(panel.background = element_rect(fill = "grey95")) +
  theme(panel.grid.major = element_line(color = "black"))
LDS_PLOT
LDS_PLOT + facet_wrap(~Population)
```

Another way to visualize this is to look at the probability of assignment by
source pouplation.

```{r dapc-region-prob, fig.width = 7, fig.height = 6}
region_summary <- region.dapc$posterior %>% 
  as.data.frame() %>% 
  as_tibble() %>% 
  add_column(Individual = indNames(dat11cc)) %>% 
  add_column(`Original Population` = as.character(strata(dat11cc)$Region)) %>% 
  group_by(`Original Population`) %>%
  mutate(n = n()) %>%
  gather(Population, Posterior, -Individual, -`Original Population`, -n) %>% 
  group_by(Population, `Original Population`, n) %>% 
  summarize(`Mean Assignment` = mean(Posterior)) %>% 
  arrange(`Mean Assignment`) %>% 
  mutate(Source = case_when(
    Population == `Original Population` ~ "Original Population",
    TRUE ~ "Other"
    )) %>%
  ungroup() %>%
  arrange(desc(n)) %>%
  mutate(`Original Population` = forcats::fct_inorder(`Original Population`)) %>%
  mutate(Population = factor(Population, levels(`Original Population`))) %>%
  mutate(`Original Population` = paste0(`Original Population`, " (", n, ")")) %>%
  mutate(`Original Population` = forcats::fct_inorder(`Original Population`))
REGION_PLOT <- region_summary %>%
  filter(n >= 10) %>%
  ggplot(aes(x = Population, y = `Mean Assignment`)) +
  geom_segment(aes(xend = Population, yend = 0)) +
  geom_point(aes(fill = Source), pch = 21) + 
  facet_wrap(~`Original Population`, ncol = 3, strip.position = "top") +
  scale_fill_grey() +
  # scale_x_discrete(position = "top") +
  scale_y_continuous(limits = c(0, 1), minor_breaks = c(seq(0, 1, by = 0.125)), breaks = c(0, 0.5, 1)) +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "top") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  # theme(axis.text.x.top = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
  theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
  # theme(panel.spacing.y = unit(0, "line")) +
  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
  theme(panel.border = element_blank()) +
  xlab("Population Assignment") +
  ylab("Mean Probability")
REGION_PLOT

```

```{r dapc-region-dualplot, fig.width = 13, fig.height = 6}
cowplot::plot_grid(REGION_PLOT, LDS_PLOT, 
                   nrow = 1, 
                   rel_widths = c(1, 0.85),
                   label_size = 16,
                   labels = "auto")
ggsave(filename = "results/figures/publication/DAPC.pdf")
```

## DAPC predictions

Since we saw MLGs crossing populations, we want to find out what happens when
we remove these MLGs, run DAPC on the regions, and then predict where they came
from.

```{r dapc-mlg, fig.width = 7, fig.height = 7}
minds <- mlg.crosspop(dat11, ~Region, indexreturn = TRUE)
dat11.nocross <- dat11[!mll(dat11) %in% minds]
dat11.nocross
dat11.nc.dapc <- dapc(dat11.nocross, strata(dat11.nocross)$Region, n.pca = 11, n.da = 14)
scatter(dat11.nc.dapc)
dat11.cross <- clonecorrect(dat11[mll(dat11) %in% minds], NA)
pred <- predict.dapc(dat11.nc.dapc, dat11.cross)
ggcompoplot(pred, setPop(dat11.cross, ~Region), cols = 2, pal = funky)
posterior <- as.data.frame(pred$posterior) %>% 
  rownames_to_column("Sample") %>%
  bind_cols(data_frame(MLG = mll(dat11.cross))) %>%
  gather(key = "population", value = "posterior", -Sample, -MLG) %>%
  group_by(Sample) %>%
  mutate(entropy = vegan::diversity(posterior)) %>%
  mutate(label = ifelse(posterior == max(posterior), MLG, NA)) %>%
  ungroup() %>%
  mutate(bins = cut(entropy, quantile(entropy)))
ggplot(posterior, aes(x = population, y = posterior, color = entropy, group = Sample)) +
  geom_line() +
  ggrepel::geom_label_repel(aes(label = label)) +
  viridis::scale_color_viridis() +
  facet_wrap(~bins, ncol = 1)
```

```{r}
top5 <- table(mll(dat11)) %>% 
  sort(decreasing = TRUE) %>% 
  head(5) %>%
  names() %>%
  as.integer()
top5
center_point <- function(n){
  fromto <- c(-1, 1)
  if (n < 4){
    fromto <- fromto/8
  } else if (n < 10){
    fromto <- fromto/4
  } else {
    fromto <- fromto/2
  }
  seq(from = fromto[1], to = fromto[2], length.out = n)
}
origins <- mlg.crosspop(dat11, ~Region, mlgsub = top5, quiet = TRUE, df = TRUE) %>%
  mutate(MLG = gsub("MLG.", "", MLG)) %>%
  mutate(population = match(as.character(Population), posterior$population %>% unique %>% sort)) %>%
  # group_by(Population) %>%
  # mutate(population = population + center_point(n())) %>%
  I()
posterior %>%
  filter(MLG %in% top5) %>%
  ggplot(aes(x = population, y = posterior, color = factor(MLG), group = MLG)) +
  geom_line() +
  geom_vline(aes(xintercept = population), data = origins, lty = 2) +
  scale_color_brewer(palette = "Dark2") +
  facet_wrap(~MLG, ncol = 1) +
  labs(list(
    lty = "Populations discovered in",
    color = "MLG"
  ))
scatter(dat11.nc.dapc)
```

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>