### 2017-07-11 doc/RMD/mlg-mcg.Rmd
add counts of MLG/MCG pairs

---
title: "Multilocus Genotype And Mycelial Compatability Group Assessment"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

# Introduction

The purpose of this document is to assess the distribution of the MLGs and MCGs
within the data. Specifically, we want to know if there are any MLGs that 
consistently coordinate with a single MCG, or if there are anything close.

## Packages and Data

```{r load_packages, message = FALSE, warning = FALSE}
library('igraph')
library('ggraph')
library('tidyverse')
library('poppr')
library('viridis')
library('visNetwork')
```

```{r load_data}
load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
load(file.path(PROJHOME, "data", "mlg-crosspop-graph.rda"))
dat11
g11o <- igraph::cluster_optimal(graph11loc$total)
comm <- igraph::communities(g11o)
names(comm) <- c("International", "Costal", "Midwest")
comm
strata(dat11) <- strata(dat11) %>%
  mutate(MLGRegion = case_when(
    .$Region %in% comm$International ~ "International",
    .$Region %in% comm$Costal ~ "Costal",
    .$Region %in% comm$Midwest ~ "Midwest",
    TRUE ~ as.character(.$Region)
  ))

setPop(dat11) <- ~MLGRegion
pal <- unlist(map2(LETTERS[2:4], comm, function(a, b) setNames(viridis::viridis(length(b), option = a), b)))
pal <- c(pal, setNames(rep("#F6F6F6FF", 3), c("Mexico", "ID", "WI")))

strat <- bind_cols(strata(dat11), 
                   other(dat11)$meta, 
                   data_frame(MLG = mll(dat11, "original")))
```

## MLG table

First, it would be nice to visualize the MLGs across populations

```{r mlg_table, fig.width = 20, fig.height = 5, keep = "last"}
mtab <- mlg.table(dat11, color = TRUE)
mplot <- last_plot()
mplot + scale_fill_viridis(discrete = TRUE, direction = -1, option = "C") +
  aes(color = I("black")) 
```

Now we can take a look at the concordance of MLGs to MCGs. We can do this by 
creating a contigency table. Of course, since we have 87 and well over 100 MLGs,
this means that the contingency table is going to be big, so to summarize it
further, I'm creating two tables, one based on MLGs that will count the number
of MCGs within each MLG and vice-versa. Of course we lose information like, if
an MCG contains several MLGs, how can we tell what the abundance is? A handy 
measure is Evenness, which scales from 0 to 1, indicating how skewed the 
observations are.

```{r}
mll.custom(dat11) <- strata(dat11)$MCG
mcgmlg <- as.data.frame(table(mll(dat11, "original"), mll(dat11, "custom"))) %>%
  setNames(c("MLG", "MCG", "Freq")) %>%
  mutate(MLG = as.character(MLG)) %>%
  mutate(MCG = as.character(MCG)) %>%
  as_tibble() %>%
  filter(Freq > 0)
mcgs <- mcgmlg %>%
  arrange(MCG) %>%
  group_by(MCG) %>%
  mutate(MLGs = sum(Freq > 0), 
         Samples = sum(Freq), 
         Evenness = diversity_stats(Freq)["E.5"], 
         data = list(data_frame(MLG = MLG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
  arrange(desc(MLGs)) %>%
  ungroup() %>%
  distinct(MCG, MLGs, Samples, Evenness, data)
mlgs <- mcgmlg %>%
  arrange(MLG) %>%
  group_by(MLG) %>%
  mutate(MCGs     = sum(Freq > 0), 
         Samples  = sum(Freq), 
         Evenness = diversity_stats(Freq)["E.5"], 
         data     = list(data_frame(MCG = MCG, Freq = Freq) %>% arrange(desc(Freq)))) %>%
  arrange(desc(Samples), desc(MCGs)) %>%
  ungroup() %>%
  distinct(MLG, MCGs, Samples, Evenness, data)
mcgs
mlgs

```

These tables are a good start, but we want to have a publication-ready table.

```{r production-tables, results = 'asis'}
fct2int <- function(x) as.integer(as.character(x))

psex_from_graph <- function(g, MLG){
  weights <- setNames(E(g)$weight, E(g)$label)
  weights <- split(weights, names(weights))
  if (length(MLG) > 1){
    weights <- weights[MLG]
    res <- 1 - vapply(weights, unique, numeric(1))
  } else {
    res <- 1 - unique(weights[[MLG]])
  }
  res
}

# Paste region and count vectors
pasteRC <- function(R, C){
  ord <- rev(order(C))
  RC <- paste(R[ord], C[ord], sep = " (")
  RC <- paste(RC, collapse = "), ")
  paste0(RC, ")")
}

mlg_table <- mcgmlg %>%
  filter(MLG %in% mlgs$MLG[1:5]) %>%
  mutate(MLG = as.integer(MLG)) %>%
  select(-Freq) %>%
  inner_join(strat) %>%
  group_by(MLG, MCG, Region) %>%
  summarize(N = n()) %>%
  ungroup() %>%
  group_by(MLG, MCG) %>%
  summarize(Region = pasteRC(Region, N), N = sum(N)) %>%
  ungroup() %>%
  mutate(mlmatch = match(MLG, mlgs$MLG[1:5])) %>%
  arrange(mlmatch, desc(N), MCG) %>%
  group_by(MLG) %>%
  mutate(Psex = psex_from_graph(graph11loc$total, as.character(unique(MLG)))) %>%
  ungroup() %>%
  select(MLG, Psex, MCG, Region) %>%
  mutate(Psex = ifelse(duplicated(MLG), "", format(Psex, scientific = FALSE, digits = 2))) %>%
  mutate(MLG = ifelse(duplicated(MLG), "", as.character(MLG)))

mcg_table <- mcgmlg %>%
  filter(MCG %in% mcgs$MCG[1:5]) %>%
  mutate(MLG = as.integer(MLG)) %>%
  select(-Freq) %>%
  inner_join(strat) %>%
  group_by(MLG, MCG, Region) %>%
  summarize(N = n()) %>%
  ungroup() %>%
  group_by(MLG, MCG) %>%
  summarize(Region = pasteRC(Region, N), N = sum(N)) %>%
  ungroup() %>%
  mutate(mlmatch = match(MCG, mcgs$MCG[1:5])) %>%
  arrange(mlmatch, desc(N), MLG) %>%
  select(MCG, MLG, Region) %>%
  mutate(MCG = ifelse(duplicated(MCG), "", as.character(MCG))) %>%
  mutate(MLG = as.character(MLG))

kable(mlg_table, caption = "Top 5 MLGs with associated MCGs and Regions (number of samples in parenthesis)")
if (!dir.exists(file.path(PROJHOME, "results/tables"))) dir.create(file.path(PROJHOME, "results/tables"))
readr::write_csv(mlg_table, path = file.path(PROJHOME, "results/tables/mlg_table.csv"))

kable(mcg_table, caption = "Top 5 MCGs with associated MLGs and Regions (number of samples in parenthesis)")
if (!dir.exists(file.path(PROJHOME, "results/tables"))) dir.create(file.path(PROJHOME, "results/tables"))
readr::write_csv(mcg_table, path = file.path(PROJHOME, "results/tables/mcg_table.csv"))
```

It might be better to visualize these data as barplots. Here we are mapping the
type (MCG/Count) to color and the opacity (alpha) to Evenness.

```{r barplots, fig.width = 10, fig.height = 5}
mcg_graph <- mcgs %>% 
  filter(Samples > 1) %>%
  gather(type, count, MLGs, Samples, -Evenness) %>%
  arrange(desc(type), desc(count)) %>%
  rename(Type = type) %>%
  mutate(MCG = forcats::fct_inorder(MCG, ordered = TRUE)) %>%
  ggplot(aes(x = MCG, y = count, group = Type, fill = Type, alpha = Evenness)) +
  geom_col(aes(width = ifelse(Type == "MLGs", 0.5, 0.85)), color = "black", position = "identity") +
  # annotate(geom = "text", x = 18, y =35, 
  #          label = sprintf("Mean Evenness: %.3f", mean(mcgs$Evenness, na.rm = TRUE))) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal(base_size = 16, base_family = "Helvetica") +
  # coord_flip() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.grid.major.y = element_line(color = "grey50")) +
  theme(panel.grid.minor.y = element_line(color = "grey70")) +
  theme(legend.box.just = "top") +
  theme(legend.box.margin = unit(c(0, 6.55, 0, 0), "lines")) +
  theme(legend.box.background = element_rect(fill = "white", color = "black")) +
  theme(legend.box = "horizontal") +
  theme(aspect.ratio = 1/2) +
  labs(list(
    alpha = "MLG\nEvenness"
  ))

even <- mcgs %>% 
  filter(MCG %in% c(53, 9)) %>% 
  mutate(MCG = ifelse(MCG == 53, "Evenness = 1.00", "Evenness = 0.55")) %>%
  unnest() %>% 
  ggplot(aes(x = forcats::fct_inorder(MLG), y = Freq, alpha = Evenness)) + 
  geom_col(color = "grey30", width = 0.75) + 
  theme_void() +
  theme(strip.text = element_text(family = "Helvetica", color = "grey30", face = "bold")) +
  theme(aspect.ratio = 0.66) +
  theme(panel.spacing = unit(0, "line")) +
  theme(legend.position = "none") +
  # theme(plot.background = element_rect(fill = NA, colour = "black")) +
  facet_wrap(~MCG, ncol = 1, scale = "free")

vp1 <- grid::viewport(width = 0.25, height = 0.4, x = 1, y = 0.975, just = c("right", "top"))
mcg_graph <- mcg_graph + theme(legend.position = c(.763, .75))

print(mcg_graph)
print(even, vp = vp1)

if (!interactive()){
  pdf(file.path(PROJHOME, "results/figures/publication/Figure1Z.pdf"), width = 10, height = 5)
  print(mcg_graph)
  print(even, vp = vp1)
  dev.off()
}
```

```{r mlg_barplots, fig.width = 4, fig.height = 10}
mlgs %>% 
  gather(type, count, MCGs, Samples, -Evenness) %>%
  arrange(desc(type), count) %>%
  mutate(MLG = forcats::fct_inorder(MLG, ordered = TRUE)) %>%
  ggplot(aes(x = MLG, y = count, group = type, fill = type)) +
  geom_col(aes(width = ifelse(type == "MCGs", 0.5, 0.85), alpha = Evenness), color = "black", position = "identity") +
  annotate(geom = "text", x = 20, y = 21, label = sprintf("Mean Evenness: %.3f", mean(mlgs$Evenness, na.rm = TRUE))) +
  # scale_fill_viridis(end = 0.75, discrete = TRUE, direction = -1) +
  scale_fill_manual(values = c("black", "white")) +
  scale_y_continuous(expand = c(0, 2)) +
  theme_minimal() +
  theme(panel.grid.major.x = element_blank()) +
  coord_flip()
  # theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  ggtitle("How Evenly are MCGs spread across Multilocus Genotypes (MLGs)?")
```

## making a graph

I believe that making a graph to visualize this might help me understand what the h\*ck is going on.

```{r}

make_mcgmlg_graph <- function(x){
  gdf <- mutate(x, MLG = paste0('MLG.', MLG))
  MLGS <- gdf %>% 
    group_by(MLG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MLG)
  MCGS <- gdf %>% 
    group_by(MCG) %>%
    summarize(size = sum(Freq)) %>%
    rename(vertex = MCG)
  VAT <- bind_rows(MLGS, MCGS)
  g <- gdf %>% 
    select(MCG, MLG, Freq) %>%
    rename(weight = Freq) %>%
    graph_from_data_frame(vertices = VAT)
  V(g)$type <- ifelse(grepl("MLG", V(g)$name), "Multilocus Genotype", "Mycelial Compatibility Group")
  g
}

add_communities <- function(g, clusters){
  comm <- communities(clusters)
  commlist <- setNames(rep(names(comm), lengths(comm)), unlist(comm, use.names = FALSE))
  V(g)$community <- commlist[V(g)$name]
  g
}
g <- make_mcgmlg_graph(mcgmlg)
g <- add_communities(g, cluster_walktrap(g))
osize <- V(g)$size
```

Because I have more control over the size and feel of the graph, I'm going to use
ggraph. Of course, since this IS a complicated data set, It's not going to be
very pretty to look at, but I'm going to save it as supplementary materials
because it's valuable to at least look this ugliness in the face and say, "Yeah,
I guess it's not so simple after all."

```{r, fig.width = 10, fig.height = 10}

V(g)$size <- sqrt(osize)/10
set.seed(2017-05-03)
lay2 <- create_layout(g, layout = "igraph", algorithm = "nicely")

the_communities <- data_frame(vertex = lay2$name, community = lay2$community) %>%
  mutate(comm = case_when(
    .$community == 7 ~ "A",
    .$community == 5 ~ "B",
    .$community == 1 ~ "C",
    TRUE ~ as.character(.$community)
  )) %>%
  group_by(community) %>%
  mutate(comm = ifelse(n() > 10, paste("Community", comm), "Other Communities (n < 10)")) %>%
  mutate(comm = comm) # It just so happens that the communities are in order by size.
lay2$community <- the_communities$comm
mcg_mlg_graph <- ggraph(lay2) +
  geom_node_circle(aes(r = size, lty = type, fill = community)) +
  geom_edge_link(aes(start_cap = circle(node1.size, unit = "native"), 
                     end_cap = circle(node2.size, unit = "native"), 
                     width = weight),
                 arrow = arrow(length = unit(0.01, "native"))) +
  geom_node_text(aes(label = gsub("MLG.", "", name), 
                     color = community, 
                     size = size/10), 
                 show.legend = FALSE) +
  coord_fixed() +
  scale_linetype_manual(values = c(1, 3)) +
  scale_edge_width(range = c(0.25, 1.5)) +
  scale_color_manual(values = c("black", "black", "white", "white")) +
  scale_fill_viridis(direction = -1, discrete = TRUE) +
  theme_graph(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "bottom", legend.direction = "vertical") +
  theme(legend.box.spacing = unit(-1, "lines")) +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines"))
  # ggtitle("Relation of Multilocus Genotypes and MCGs")

mcg_mlg_graph
ggsave(plot = mcg_mlg_graph, filename = file.path(PROJHOME, "results/figures/publication/FigureS2.pdf"), 
       width = 88*3, height = 88*3.5, units = "mm")
```

Of interest is accounting for the communities that have a single MCG/MLG pair.
We can detect these by filtering for communities that only have two members:

```{r}
singles <- the_communities %>% 
  count(community) %>% 
  arrange(desc(n)) %>% 
  filter(n == 2) %>% 
  inner_join(the_communities, by = "community") %>% 
  mutate(type = case_when(grepl("MLG", vertex) ~ "MLG", TRUE ~ "MCG")) %>% 
  spread(type, vertex) %>%
  select(-n, -comm)
# assessing that this truely contains the singletons
nrow(inner_join(singles, the_communities, by = c("MCG" = "vertex")))
nrow(inner_join(singles, the_communities, by = c("MLG" = "vertex")))

knitr::kable(singles)
```

### Interactive Visualizations

The *visNetwork* package wraps a java library that has very good network
visualization routines. I can conver this graph to that.

```{r, eval = FALSE}

make_rgb <- function(x, alpha = 1){
  out <- col2rgb(x, alpha = TRUE)
  out[4, ] <- floor(out[4, ]*alpha)
  out      <- apply(out, 2, paste, collapse = ", ")
  paste0("rgba(", out, ")")
}
vg <- g %>%
  set_vertex_attr("size", value = osize) %>%
  set_vertex_attr("value", value = osize) %>%
  set_vertex_attr("color", value = viridis(4)[as.integer(factor(the_communities$comm))]) %>%
  set_vertex_attr("shape", value = ifelse(V(.)$type == "Multilocus Genotype", "triangle", "circle")) %>%
  set_edge_attr("width", value = E(.)$weight) %>%
  toVisNetworkData()
vg$nodes <- vg$nodes %>% 
  group_by(id) %>%
  mutate(color = list(list(background = make_rgb(color, 0.8), 
                      border = make_rgb(rep("black", n()), 0.8),
                      highlight = list(background = make_rgb(color),
                                       border = make_rgb(rep("black", n()))
                                       )
                      )))
vgn <- visNetwork(nodes = vg$nodes, edges = vg$edges,# height = "500px", 
                  main = "Relation of Multilocus Genotypes and MCGs")
set.seed(2017-05-03)
vgn
# vgn %>%
#   visIgraphLayout("layout_nicely") %>%
#   visOptions(highlightNearest = list(enabled = TRUE, hover = TRUE), 
#              nodesIdSelection = TRUE)
#   
```

## Subgraph

So that's a big hairy graph. What happens when we look at a subgraph of the
top 5 MCGs?

```{r top5graph, fig.height = 7, fig.width = 7}
top5 <- filter(mcgmlg, as.character(MCG) %in% mcgs$MCG[1:5])
top5g <- make_mcgmlg_graph(top5)
tosize <- V(top5g)$size
V(top5g)$size <- sqrt(tosize)/10
set.seed(2017-05-08)
top5lay <- create_layout(top5g, layout = "igraph", algorithm = "nicely")
top5lay$community <- inner_join(data_frame(vertex = top5lay$name), the_communities)$comm
t5g <- mcg_mlg_graph %+% top5lay + ggtitle("Top 5 Mycelial Compatibility Groups and associated MLGs")
t5g
ggsave(t5g, file = file.path("results", "figures", "publication", "FigureS2B.pdf"), width = 88, height = 88, scale = 3, units = "mm")
```

Vey nice!

There are a whole buttload of MLGs for those 5 MCGs. What are the severity 
ratings for those? Again, we take take a look at these AND simultaneously query
the top 5 MLGs for this.

```{r severity, fig.width = 7, fig.height = 4}
count_group <- . %>% 
  mutate(nobs = n()) %>%  # count the number of samples/MCG
  ungroup() %>%
  arrange(desc(nobs)) # arrange by number of samples and reorder factors 
Severity <- filter(strat, MCG %in% mcgs$MCG[1:5]) %>%
  group_by(MCG) %>% 
  count_group %>%
  mutate(MCG = forcats::fct_inorder(factor(MCG), ordered = TRUE)) %>%
  mutate(MLG = ifelse(MLG %in% mlgs$MLG[1:5], paste("MLG", MLG), "Other"))

severity_plot <- ggplot(Severity, aes(x = MCG, y = Severity)) +
  geom_point(aes(fill = MLG), 
             position = position_jitter(width = 0.2),
             alpha = 0.75,
             pch = 21) +
  scale_fill_viridis(discrete = TRUE, direction = -1) +
  theme_bw() +
  theme(legend.position = "bottom") +
  theme(aspect.ratio = 0.6) +
  facet_wrap(~Region, nrow = 2) +
  ylim(c(3.5, 8)) +
  labs(list(
    title = "Severity by MCG and Region",
    fill = "Multilocus Genotype",
    subtitle = "Five most abundant multilocus genotypes shown"
  ))
severity_plot
ggsave(severity_plot, filename = "results/figures/publication/FigureS3.pdf", width = 183, unit = "mm")
```

## Assessing MCG/MLG community structure/cohesiveness

The communities defined above are defined by the connections between different
MCGs. There is still the question of how cohesive these communities are. I 
believe one way to investigate this is to add the communities assignments back
to the data and assess how well they split up on a dendrogram created with 
genetic distance.

```{r treebuilding, cache = TRUE}
strat_mcg <- left_join(strat, the_communities, by = c("MCG" = "vertex"))
datdist   <- bruvo.dist(dat11, replen = other(dat)$REPLEN)
datree    <- phangorn::upgma(datdist)
```

Now we can visualize that trees.

```{r trees, fig.width = 7, fig.height = 7}
ape::plot.phylo(ape::ladderize(datree), type = "fan", no.margin = FALSE, tip.col = viridis(4, direction = -1)[as.integer(factor(strat_mcg$comm))])
legend(-0.015, 0.11, legend = levels(factor(strat_mcg$comm)), fill = viridis(4, direction = -1), bty = "n")
title("MCG-based assignment")

ape::plot.phylo(ape::ladderize(datree), type = "fan", no.margin = FALSE, tip.col = plasma(366)[rank(strat$Severity, ties.method = "first")])
legend(-0.015, 0.125, legend = quantile(strat$Severity), fill = plasma(5), bty = "n")
title("Severity")
```

From what I can see, it appears that there's not much of a tight correlation of 
genetic structure and MCG. There is the clear cluster of MLG 66 in Community C,
but there are other isolates in distant parts of the tree. I'm wondering what
happens if we average the distances between groups.

```{r group_dist}
datmat <- as.matrix(datdist)
distmcg <- strat_mcg %>% 
  group_by(comm) %>%
  summarize(dist = list(as.dist(datmat[Isolate, Isolate])), N = n())
distmcg %>%
  group_by(comm) %>%
  mutate(meandist = mean(dist[[1]]),
         mediandist = median(dist[[1]])) %>%
  ungroup() %>%
  select(comm, N, meandist, mediandist)
distmcg %>% 
  unnest() %>%
  ggplot(aes(x = dist, fill = comm)) +
  geom_density(alpha = 0.5) +
  geom_rug(aes(color = comm), alpha = 0.12) +
  scale_fill_viridis(direction = -1, discrete = TRUE) +
  scale_color_viridis(direction = -1, discrete = TRUE) +
  ggtitle("Bruvo's distance by MCG community") +
  theme_bw(base_size = 14)
```

This is quite distant. For reference, a distance of `r round((0.5/11)*7, 2)` is
on average 7 steps.

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>