### 2017-07-26 doc/RMD/MLG-distribution.Rmd
add colnames parameter

---
title: "Analyzing MLG distribution"
output: 
  html_notebook:
    toc: true
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

In this document, I will create a graph that shows the distribution of MLGs
across populations

```{r load_packages, message = FALSE, warning = FALSE}
library('tidyverse')
library('poppr')
library('igraph')
library("ggraph")
```

## Loading data and setting strata

I'm loading the data and splitting it into the 11 locus and 16 locus versions

```{r}
load("data/sclerotinia_16_loci.rda")

```

Now we can inspect the genotype accumulation curves. Notice that the 16 loci
version plateaus, but we know that these extra loci are inconsistent, so we 
might not want to trust them.

```{r genotype_curves, fig.width = 5, fig.height = 5}
locus_table(dat)
set.seed(2017-06-28)
genotype_curve(dat, sample = 1000, quiet = TRUE)
genotype_curve(dat11, quiet = TRUE)
```

Creating contracted genotypes for use later

```{r}
mlg.filter(dat, distance = bruvo.dist, replen = other(dat)$REPLEN) <- .Machine$double.eps
dat
mlg.filter(dat11, distance = bruvo.dist, replen = other(dat11)$REPLEN) <- .Machine$double.eps
dat11
mll(dat) <- "original"
mll(dat11) <- "original"
```

## Crossing populations

We can use `mlg.crosspop()` to tabulte which MLGs cross populations.

I realized that it's possible to use an MLG table with matrix multiplication to
get an adjency matrix. Here I'm making a function to create a list of graphs.

The vertices, representing populations, will have the following attributes:

 - Size: equal to the number of MLGs within the population
 - Weight: fraction of private MLGs

The edges, representing multilocus genotypes are a bit trickier to handle since
multiple edges can represent a single MLG passing through populations. They have
the following attributes:

 - Label: the MLG this edge belongs to. Note that multiple edges can share the
          same label.
 - Width: the number of populations this MLG passes through
 - Weight: 1 - (probability of a second encounter by chance); 1 - psex

The reason why I'm weighting these nodes as 1 - psex is for subsequent community
analysis. The more likely you are to encounter a genotype by chance, the less
influence that genotype should have on the connectivity.

```{r}
rerange <- function(x){
  minx <- min(x, na.rm = TRUE)
  maxx <- max(x, na.rm = TRUE)
  if (minx < 0)
    x <- x + abs(minx)
    maxx <- maxx + abs(minx)
  if (maxx > 1)
    x <- x/maxx
  return(x)
}
make_graph_list <- function(dat){ # dat is a genclone/snpclone object
  # w       <- pgen(dat, by_pop = FALSE) %>% rowSums() %>% exp() %>% setNames(paste0("MLG.", mll(dat)))
  # w       <- w[unique(names(w))]
  w       <- psex(dat, by_pop = FALSE, method = "multiple") %>% split(mll(dat))
  names(w)<- paste0("MLG.", names(w))
  datmlg  <- mlg.table(dat, plot = FALSE) > 0 # presence/absence of MLG
  crosses <- mlg.crosspop(dat, quiet = TRUE, df = TRUE) %>% tbl_df()
  adjmat  <- datmlg %*% t(datmlg) 
  cols    <- sort(colnames(adjmat))
  adjmat  <- adjmat[cols, cols]
  # Creating Graph
  g           <- graph_from_adjacency_matrix(adjmat, mode = "undirected", diag = FALSE)
  V(g)$size   <- diag(adjmat)
  g           <- delete_vertices(g, degree(g) == 0)
  shared_mlg  <- (crosses %>% group_by(Population) %>% summarize(n = n()))$n
  V(g)$weight <- shared_mlg/V(g)$size # fraction of shared MLGs
  el          <- as_adj_edge_list(g)
  el          <- el[lengths(el) > 0]
  popgraphs <- setNames(vector(mode = "list", length = length(el) + 1), c(names(el), "total"))
  for (v in names(el)){
    idx  <- el[[v]]
    mlgs <- crosses %>%           # How to get all MLGs from a single population:
      filter(Population == v) %>%         # Grab only the population e and then
      select(MLG) %>%                     # remove everything but the MLGs to do an
      inner_join(crosses, by = "MLG") %>% # inner join of the original list and then
      filter(Population != v) %>%         # remove the query population to give
      arrange(Population)                 # the neigboring populations in order.
    MLGS <- as.character(mlgs$MLG)
    E(g)[idx]$label  <- substr(MLGS, 5, nchar(MLGS))
    E(g)[idx]$width  <- as.integer(table(MLGS)[MLGS]) # size == n populations visited
    E(g)[idx]$weight <- 1 - map_dbl(w[MLGS], 2)       # weight == 1 - psex 2nd encounter
    popgraphs[[v]]   <- subgraph.edges(g, eids = idx)
  }
  popgraphs[["total"]] <- g
  popgraphs
}

plot_mlg_graph <- function(g, glayout = NULL, label = TRUE, seed = 2017-06-28){
  set.seed(seed)
  if (is.null(glayout)){
    glayout <- layout_nicely(g)
    colnames(glayout) <- c("x", "y")
  } else {
    glayout <- glayout[V(g)$name, ]
  }
  shared_mlg  <- V(g)$weight * V(g)$size
  g           <- add_vertices(g, length(V(g)), size = V(g)$size - shared_mlg, color = "grey90")
  g2          <- g
  V(g2)$label <- ifelse(!is.na(V(g2)$name), sprintf("%s [(%d/%d)]", V(g2)$name, shared_mlg, V(g2)$size), NA)
  glay        <- create_layout(g2, "manual", node.positions = as.data.frame(rbind(glayout, glayout)))
  x_nudge     <- ifelse(abs(glay$x) == 1, -glay$x/10, glay$x/10)
  breaks      <- quantile(1 - E(g)$weight)
  breaks      <- setNames(breaks, format(breaks, digits = 2))
  outg <- ggraph(glay) +
    geom_edge_fan(aes(alpha = 1 - weight, width = width + 1)) +
    geom_node_circle(aes(r = drop(scale(size, center = FALSE))/10, fill = size))
  outg <- if (label) outg + geom_node_label(aes(label = label), repel = TRUE, parse = TRUE, label.size = 0.75, nudge_x = x_nudge) else outg
  outg + 
    viridis::scale_fill_viridis(option = "C") +
    scale_edge_alpha_continuous(range = c(1, 0.25), breaks = rev(breaks)) +
    scale_edge_width_continuous(range = c(0.25, 1.25), breaks = c(2:5)) +
    coord_fixed() +
    theme_void() +
    theme(text = element_text(size = 14)) +
    labs(list(
      title = "Shared haplotypes across regions",
      fill = "Number of\nhaplotypes",
      edge_alpha = "Probability of\nsecond encounter",
      edge_width = "Populations\nper haplotype",
      caption = "Outer circle: Number of haplotypes in the region\nInner Circle: Number of private haplotypes in the region"
    ))
}

plot_mlg_subgraph <- function(graphlist){
  for (i in names(graphlist)){
    pg   <- graphlist[[i]]
    labs <- ifelse(E(pg)$width > 1, E(pg)$label, NA)
    labs <- ifelse(duplicated(labs), NA, labs)
    plot(pg, 
         main = i, 
         layout = layout_as_star(pg, center = i), 
         edge.width = E(pg)$width,
         edge.color = grey(rerange(1 - E(pg)$weight)), 
         edge.label = labs)
  }
}

good_layout <- read.table(
text = 
"   x                            y
Australia  0.8090169943749470071737  5.877852522924730260812e-01
CA -1.0000000000000000000000  1.224646799147350002426e-16
CO -0.3090169943749480063744 -9.510565162951539752711e-01
France -0.3090169943749470071737  9.510565162951539752711e-01
MI  0.0000000000000000000000  0.000000000000000000000e+00
MN  1.0000000000000000000000  0.000000000000000000000e+00
ND  0.3090169943749470071737 -9.510565162951539752711e-01
NE -0.8090169943749470071737 -5.877852522924730260812e-01
NY  0.8090169943749470071737 -5.877852522924730260812e-01
OR  0.3090169943749470071737  9.510565162951539752711e-01
WA -0.8090169943749470071737  5.877852522924730260812e-01"
) %>% as.matrix()

alt_layout <- read.table(
text = 
"   x                            y
Australia  0.8090169943749470071737  5.877852522924730260812e-01
OR -1.0000000000000000000000  1.224646799147350002426e-16
CO -0.3090169943749480063744 -9.510565162951539752711e-01
WA -0.3090169943749470071737  9.510565162951539752711e-01
MI  0.0000000000000000000000  0.000000000000000000000e+00
MN  1.0000000000000000000000  0.000000000000000000000e+00
ND  0.3090169943749470071737 -9.510565162951539752711e-01
NE -0.8090169943749470071737 -5.877852522924730260812e-01
NY  0.8090169943749470071737 -5.877852522924730260812e-01
France  0.3090169943749470071737  9.510565162951539752711e-01
CA -0.8090169943749470071737  5.877852522924730260812e-01"
) %>% as.matrix()
alt_layout <- alt_layout[rownames(good_layout), ]

```

## Graphs

With the fuctions above, we can create and plot the graphs.

```{r}
# Creating the graphs
graph16loc <- make_graph_list(dat)
graph11loc <- make_graph_list(dat[loc = keeploci, mlg.reset = TRUE])
```

### Individual population subgraphs

Each subgraph shows all of the connections for a single population.

```{r subgraphs, fig.width = 10, fig.height = 10}
# Plotting the subgraphs
par(mfrow = c(3, 4))
plot_mlg_subgraph(graph16loc[-length(graph16loc)])

par(mfrow = c(3, 4))
plot_mlg_subgraph(graph11loc[-length(graph11loc)])
```

### Cross-regional graphs

I'm plotting two graphs here because it's useful to see if both of them are
congruent in their community predictions.

#### With 16 loci

First, I'm going to show the results of a community analysis. I'm using the 
igraph function `cluster_optimal()` to cluster the nodes.

```{r graph-16-loci, fig.width = 7, fig.height = 7}
par(mfrow = c(1, 1))
(g16o <- cluster_optimal(graph16loc$total))
plot_mlg_graph(graph16loc$total, alt_layout) + labs(list(subtitle = "(16 loci)"))
ggsave(filename = "results/figures/publication/FigureS1Z.pdf", width = 88, height = 88, units = "mm", scale = 2)
```

What we see is that We are given three clusters showing a clustering of the
plains states, the west coast, and Australia, France, and Minnesota. The last
cluster is likely driven by the single genotype shared between these three
regions that has a low probability of a second encounter.

#### With 11 loci

```{r graph-11-loci, fig.width = 7, fig.height = 7}
(g11o <- cluster_optimal(graph11loc$total))
plot_mlg_graph(graph11loc$total, alt_layout) + labs(list(subtitle = "(11 loci)"))
ggsave(filename = "results/figures/publication/Figure3Z.pdf", width = 88, height = 88, units = "mm", scale = 2)
```

```{r, saving_data}
save(graph16loc, graph11loc, alt_layout, file = "data/mlg-crosspop-graph.rda")
```

### Region Cohesion

Here I am using DAPC to assess how cohesive these artificial regions are.

```{r, region_cohesion}
g11o <- igraph::cluster_optimal(graph11loc$total)
comm <- igraph::communities(g11o)
names(comm) <- c("International", "Costal", "Midwest")
comm
strata(dat11) <- strata(dat11) %>%
  mutate(MLGRegion = case_when(
    .$Region %in% comm$International ~ "International",
    .$Region %in% comm$Costal ~ "Costal",
    .$Region %in% comm$Midwest ~ "Midwest",
    TRUE ~ as.character(.$Region)
  ))

setPop(dat11) <- ~MLGRegion
```

```{r region_dapc}
set.seed(2017-07-12)
reg <- xvalDapc(tab(dat11), pop(dat11), n.rep = 100)
reg$DAPC
scatter(reg$DAPC, col = viridis::viridis(6), bg = "grey90")
ggcompoplot::ggcompoplot(reg$DAPC, dat11, pal = viridis::viridis(6), cols = 2)
reg$DAPC$posterior %>%
  as_tibble() %>%
  bind_cols(data.frame(Population = reg$DAPC$grp)) %>%
  group_by(Population) %>%
  summarize_all(mean) %>%
  gather(assignment, probability, -Population) %>%
  ggplot(aes(x = assignment, y = probability)) +
  geom_segment(aes(xend = assignment, yend = 0)) +
  geom_point(pch = 21, fill = "grey") +
  facet_wrap(~Population) +
  scale_y_continuous(limits = c(0, 1), 
                     minor_breaks = c(seq(0, 1, by = 0.125)), 
                     breaks = c(0, 0.5, 1)) +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(legend.position = "top") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  theme(axis.ticks.y = element_blank()) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
  theme(panel.border = element_blank()) +
  xlab("Population Assignment") +
  ylab("Mean Probability")
```

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>