### 2017-08-01 doc/RMD/MCG-virulence.Rmd
Add aggressiveness tests to fix #8

---
title: "Assessing Virulence by MCG"
output: 
  html_notebook:
    toc: true
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

```{r load_packages, message = FALSE, warning = FALSE}
library("tidyverse")
```

# Purpose

This document is to assess virulence associated with the 10 most common MCGs. We
can get this from the cleaned data set that we saved in the
`data-comparison.Rmd` file. Note, we will be labelling the plot with 
"Aggressiveness" as that is the preferred term.

```{r load_data}
datcols <- cols(
  .default = col_integer(),
  Severity = col_double(),
  Region = col_character(),
  Source = col_character(),
  Host = col_character()
)
dat <- read_csv(file.path(PROJHOME, "data", "clean_data.csv"), col_types = datcols) %>%
  select(Severity, MCG, Region, Source)
dat
```

Now, I can filter out the top 10 MCGs:

```{r filtering}
top_mcg <- dat %>% 
  group_by(MCG) %>%
  summarize(N = n()) %>%
  arrange(desc(N)) %>%
  slice(1:10) %>% 
  inner_join(dat, by = "MCG") %>%
  select(MCG, Severity) %>%
  mutate(MCG = forcats::fct_inorder(as.character(MCG)))
top_mcg
```

## Visualizing distributions

```{r vis, fig.width = 7, fig.height = 4}
set.seed(2017-06-29)
ggplot(top_mcg, aes(x = MCG, y = Severity)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.25), alpha = 0.5) +
  scale_y_continuous(limits = c(1, 9), breaks = c(1, 3, 5, 7, 9), expand = c(0, 0.1)) +
  # scale_x_discrete(position = "top") +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  # theme(axis.ticks.x = element_blank()) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(colour = "grey50")) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank()) +
  labs(list(
    # title = "Aggressiveness for the top 10 MCGs",
    y = "Aggressiveness",
    x = "Mycelial Compatibility Group"
    ))
```

```{r virulence_table}
top_mcg %>% 
  group_by(MCG) %>%
  summarize(N = n(), 
            `Min Aggressiveness` = min(Severity),
            `Max Aggressiveness` = max(Severity),
            `Average Aggressiveness` = mean(Severity)
            ) %>%
  knitr::kable(digits = 2)
```

## ANOVA

The default ANOVA in R sets contrasts as `contrast.treatment`, which compares
everything to the first factor, considered the treatment. Since we are
interested in whether or not there IS a difference between samples, this will
be sufficient.

After the ANOVA, we performed a Tukey's Honest Significant Difference test to
see exactly what groups these fell into.

```{r anova, fig.width = 4, fig.height = 8}
ANOVA <- aov(Severity ~ MCG, data = top_mcg)
ANOVA
summary(ANOVA)
plot(TukeyHSD(ANOVA, conf.level = 0.95), las = 2)
grps <- agricolae::HSD.test(ANOVA, "MCG", alpha = 0.05)$groups
readr::write_csv(grps, path = "results/tables/MCG-aggressiveness.csv", col_names = TRUE)

grps %>% 
  dplyr::rename(MCG = trt) %>%
  dplyr::rename(Group = M) %>%
  dplyr::rename(`Mean Aggressiveness` = means) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
  huxtable::set_col_width(c(0.05, 0.3, 0.08)) %>%
  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
  huxtable::print_md(max_width = 31)
```

There appears to be a significant effect at p = `r formatC(broom::tidy(ANOVA)$p.value[1])`.

# Assessing Virulence by Region

Since this was done in Otto-Hanson *et al.*, it would be a good idea for us to 
assess this as well.

```{r fig.width = 5, fig.height = 3.75}
set.seed(2017-06-29)
region_dat <- dat %>%
  group_by(Region) %>%
  filter(n() > 5) %>%
  mutate(mean_sev = mean(Severity)) %>%
  mutate(Source = ifelse(Source == "wmn", "wmn", "other")) %>%
  arrange(desc(mean_sev)) %>%
  ungroup() %>%
  mutate(Region = forcats::fct_inorder(Region)) %>%
  select(Region, Severity, Source)
ggplot(region_dat, aes(x = Region, y = Severity)) +
  geom_point(aes(fill = Source), 
             pch = 21, 
             position = position_jitter(width = 0.25), 
             alpha = 0.35) +
  geom_boxplot(alpha = 0.5) +
  scale_y_continuous(limits = c(1, 9), breaks = c(1, 3, 5, 7, 9), expand = c(0, 0.1)) +
  scale_fill_manual(values = rev(c("black", "white"))) +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(colour = "grey50")) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank()) +
  theme(legend.position = "top") +
  theme(plot.margin = unit(c(0, 0, 0, 0), "lines")) +
  labs(list(
    y = "Aggressiveness",
    x = "Region"
    ))
if (!interactive()){
  ggsave(file.path("results", "figures", "publication", "aggressiveness.pdf"),
         width = 88, units = "mm")
}
```

```{r anova2, fig.width = 4, fig.height = 8}
region_dat %>%
  group_by(Region) %>%
  summarize(N = n(), 
            `Min Aggressiveness` = min(Severity),
            `Max Aggressiveness` = max(Severity),
            `Average Aggressiveness` = mean(Severity),
            `Median Aggressiveness` = median(Severity)
            ) %>%
  knitr::kable(digits = 2)
ANOVA <- aov(Severity ~ Region, data = region_dat)
ANOVA
summary(ANOVA)
plot(TukeyHSD(ANOVA, conf.level = 0.95), las = 2)
grps <- agricolae::HSD.test(ANOVA, "Region", alpha = 0.05)$groups %>% 
  dplyr::rename(Region = trt) %>%
  dplyr::rename(Group = M) %>%
  dplyr::rename(`Mean Aggressiveness` = means)

readr::write_csv(grps, path = "results/tables/Region-aggressiveness.csv", col_names = TRUE)

grps %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
  huxtable::print_md(max_width = 40)
```

```{r fig.width = 5, fig.height = 3.5}
library("poppr")
load(file.path("data", "sclerotinia_16_loci.rda"))
strat <- strata(dat11) %>% 
  add_column(MLG = mll(dat11)) %>%
  add_column(Severity = other(dat11)$meta$Severity)
top_mlg <- strat %>% 
  group_by(MLG) %>%
  summarize(N = n()) %>%
  arrange(desc(N)) %>%
  slice(1:10) %>% 
  inner_join(strat, by = "MLG") %>%
  select(MLG, Severity) %>%
  mutate(MLG = forcats::fct_inorder(as.character(MLG)))
set.seed(2017-06-29)
ggplot(top_mlg, aes(x = MLG, y = Severity)) +
  geom_boxplot() +
  geom_point(position = position_jitter(width = 0.25), alpha = 0.5) +
  scale_y_continuous(limits = c(1, 9), breaks = c(1, 3, 5, 7, 9), expand = c(0, 0.1)) +
  # scale_x_discrete(position = "top") +
  theme_bw(base_size = 16, base_family = "Helvetica") +
  theme(aspect.ratio = 1/2) +
  theme(axis.text = element_text(color = "black")) +
  # theme(axis.ticks.x = element_blank()) +
  theme(panel.grid.major = element_line(colour = "grey20")) +
  theme(panel.grid.minor = element_line(colour = "grey50")) +
  theme(panel.grid.major.x = element_blank()) +
  theme(panel.border = element_blank()) +
  labs(list(
    # title = "Aggressiveness for the top 10 MCGs",
    y = "Aggressiveness",
    x = "Mulitlocus Genotype"
    ))
```

```{r anova3, fig.width = 4, fig.height = 8}
top_mlg %>%
  group_by(MLG) %>%
  summarize(N = n(), 
            `Min Aggressiveness` = min(Severity),
            `Max Aggressiveness` = max(Severity),
            `Average Aggressiveness` = mean(Severity)
            ) %>%
  knitr::kable(digits = 2)
ANOVA <- aov(Severity ~ MLG, data = top_mlg)
ANOVA
summary(ANOVA)
plot(TukeyHSD(ANOVA, conf.level = 0.95), las = 2)
grps <- agricolae::HSD.test(ANOVA, "MLG", alpha = 0.05)$groups %>% 
  dplyr::rename(MLG = trt) %>%
  dplyr::rename(Group = M) %>%
  dplyr::rename(`Mean Aggressiveness` = means)

readr::write_csv(grps, path = "results/tables/MLG-aggressiveness.csv", col_names = TRUE)

grps %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>%
  huxtable::set_number_format(row = huxtable::everywhere, col = 1, 0) %>% 
  huxtable::set_col_width(c(0.13, 0.25, 0.08)) %>%
  huxtable::set_align(row = huxtable::everywhere, col = 2, "right") %>%
  huxtable::print_md(max_width = 40)
```

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>