### 2017-08-01 doc/RMD/table-1.Rmd
change table 1 to be more concise
I changed table 1 so that it highlights the salient features of
our data.

---
title: "Table 1"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```
# Creation of Table 1

Table one is one of those long, descriptive tables that requires a lot of
fiddling. The *huxtable* package makes working with these tables slightly less
of a pain in the neck. Because this table involves wrapping and the text
can get split up, I am manually copying and pasting the result into my document.

```{r load_packages, warning = FALSE, message = FALSE}
library("readr")
library("dplyr")
library("huxtable")
```

```{r isolate_table, comment = NA}
clmn <- cols(
  .default = col_integer(),
  Severity = col_double(),
  Region = col_character(),
  Source = col_character(),
  Host = col_character()
)

# Helper function to generate positions of blanks in a column for pretty printing
blank_it <- function(x){
  res <- rep(TRUE, length(x))
  res[1] <- FALSE
  res
}

dat <- readr::read_csv(file.path(PROJHOME, "data/clean_data.csv"), col_types = clmn) %>%
  mutate(Country = ifelse(nchar(Region) == 2, "USA", Region)) %>% # Defining country column
  mutate(Region = ifelse(nchar(Region) == 2, Region, "-")) %>%    # Adding blanks for international "states"
  mutate(Source = ifelse(Source == "wmn", "wmn", "producer")) %>% # Changing fields to binary counts
  select(Country, Region, Source, Year, Host) %>%                 # Summarizing the number of isolates
  group_by_all() %>%                                              #
  summarize(N = n()) %>%                                          #
  arrange(desc(Country), Region, Source, Year) %>%                # Rearranging rows
  group_by(Country, Region, Source) %>%                           # Collapsing Years and Hosts
  summarize(Year = paste(sort(unique(Year)), collapse = ", "),    #
            Host = paste(sort(unique(Host)), collapse = ", "),    #
            N = sum(N)) %>%                                       #
  arrange(desc(Country), Region, desc(N)) %>%                     # Rearranging rows
  group_by(Country, Region) %>%                                   # Adding blanks in repeated Country and Region Names
  mutate(set_blank = blank_it(Region)) %>%                        #
  ungroup() %>%                                                   #
  mutate(Country = ifelse(set_blank, "", Country)) %>%            #
  mutate(Region = ifelse(set_blank, "", Region)) %>%              #
  select(Country, Region, Source, Year, Host, N) %>%              #
  rename(State = Region) %>%                                      # Renaming columns
  rename(`Field Code` = Source)                                   #

dt <- huxtable::as_huxtable(dat, add_colnames = TRUE) %>% # Creating the huxtable
  set_align(everywhere, col = 4, value = "right") %>%     # Aligning numeric columns to the right
  set_align(everywhere, col = 6, value = "right") %>%     # 
  set_wrap(TRUE) %>%                                      # Wrapping columns
  set_number_format(value = 0)                            # Print numbers without decimals
col_width(dt) <- c(0.055, 0.035, 0.06, 0.1, 0.151, 0.025) # Specify width for each column
print_md(dt)                                              # print in pandoc markdown format
```

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>