### 2017-08-03 doc/RMD/tree.Rmd
add tree

---
title: "tree"
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
knitr::opts_chunk$set(fig.align = "center")
```

## Packages and Data

```{r load_packages, warning = FALSE, message = FALSE}
library('tidyverse')
library('poppr')
library('ggtree')
```

```{r load_data}
load(file.path(PROJHOME, "data", "sclerotinia_16_loci.rda"))
setPop(dat11) <- ~Host/Source/Region/Year
dat11cc <- clonecorrect(dat11, ~Host/Source/Region/Year, keep = 1:4)
dat11cc

# Asserting that nothing messed up with the metadata.
stopifnot(identical(indNames(dat11cc), other(dat11cc)$meta$Isolate))
```

The purpose of this document is simply to calculate a bootstrapped tree for 
Bruvo's distance.

```{r bdtree, cache = TRUE}
set.seed(2017-08-03)
bt <- bruvo.boot(clonecorrect(dat11, strata = NA), 
                 replen = other(dat11)$REPLEN, 
                 sample = 1000, 
                 tree = "nj", 
                 showtree = FALSE)
```

Let's take a look at the results of the bootstrap analysis. Note, that we would
traditionally ignore results < 75.

```{r bsresults}
bt$node.labels
summary(bt$node.labels)
hist(bt$node.labels)
```

It's clear that our results show that NONE of the clades are well supported
(except for the whole tree, which is default), so let's take a look at how the
tree looks. One of the things we want to do is see where the popualtions fit.
Since we created a clone-corrected tree here (for speed), we are going to create
a matrix to tally up the samples from different populations per MLG.

An important thing to keep track of is the fact that ggtree works from rownames
for tip labels, so we have to add them in.

```{r}
data <- bind_cols(strata(dat11), other(dat11)$meta) %>% 
  add_column(MLG = mll(dat11))

otherdf <- data %>% 
  group_by(MLG, Region) %>%
  summarize(N = n()) %>%
  spread(Region, N)
otherdf <- inner_join(data %>% 
                        group_by(MLG) %>% 
                        summarize(N = n(), id = Isolate[1]), 
                      otherdf, 
                      by = "MLG")
df <- otherdf %>% 
  select(-MLG, -N) %>% 
  as.data.frame() %>% 
  column_to_rownames("id")
otherdf
```

Now we can create the tree with the matrix, coloring by number of isolates.

```{r the_tree, fig.width = 7, fig.height = 12}
gbt <- ggtree(bt) + geom_tippoint() + theme_tree2() + xlab("Bruvo's Distance (11 loci)")
gheatmap(gbt, df, colnames = FALSE, width = 0.6) %>% 
  scale_x_ggtree() + 
  viridis::scale_fill_viridis(guide = "legend") +
  scale_y_continuous(expand = c(0, 0)) +
  theme(text = element_text(size = 16, family = "Helvetica")) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, color = "black")) +
  theme(panel.grid.major.x = element_line(colour = "grey50", linetype = 3)) +
  labs(list(fill = "N isolates"))
```

Of course, because of the low bootstrap support, we don't have very much faith
in this tree. One of the promising aspects, however is the fact that all the
isolates from Mexico group together, as was shown in the DAPC. Moreover, we can
see that this is not driven by private alleles:

```{r private allelels}
private_alleles(dat11, locus ~ Region, count.alleles = FALSE)
```

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>