### 2017-08-17 doc/RMD/by-year.Rmd
update figures and rerun nearly everything

---
title: "Differentiation By Year"
output: 
  html_notebook:
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = PROJHOME)
```

# Purpose

Because we have collections of isolates from the same WMN across different years,
we can investigate how populations change over the years. I'm attempting to be
careful with this analysis because, much like the white mold screening nursery
analysis, there are small sample sizes, which could bias the results.

# Data

```{r packages}
library("poppr")
library("tidyverse")
library("cowplot")
make_amova_table <- function(am, amt, samples = "Region"){
  tot <- nrow(am$results)
  res <- data.frame(list(am$results[-tot, c("Df", "Sum Sq")], 
                         Percent = am$componentsofcovariance[-tot, 2],
                         Pval    = rev(amt$pvalue), 
                         Phi     = rev(am$statphi$Phi[-tot])))
  res <- as.matrix(res)
  colnames(res) <- c("d.f.", "Sum of Squares", "Percent variation", "P", 
                     "Phi statistic")
  names(dimnames(res)) <- c("levels", "statistic")
  rownames(res) <- gsub("samples", samples, rownames(res))
  return(res)
}

make_amova_printable <- function(amtab, amtabcc){
  am_array <- array(dim      = c(dim(amtab), 2),
                    dimnames = c(dimnames(amtab), 
                                 list(c("full", "clone-corrected"))))
  am_array[, , 1] <- amtab
  am_array[, , 2] <- amtabcc
  tabfun <- function(x){
    x <- paste0(paste0(signif(x, 3), collapse = " ("), ")")
    return(x)
  }
  res <- apply(am_array, c(1, 2), tabfun)
  return(res)
}
```

```{r data}
load("data/sclerotinia_16_loci.rda")
dat11cc <- dat11 %>%
  clonecorrect(~Region/Source/Host/Year) %>% 
  setPop(~Region/Year)

addStrata(dat11cc) <- strata(dat11cc) %>% 
  mutate(SourceType = forcats::fct_inorder(ifelse(Source == "wmn", "wmn", "other"))) %>%
  select(SourceType)
wmncc   <- dat11cc %>% 
  setPop(~Source) %>% 
  popsub("wmn") %>% 
  setPop(~Region/Year)
table(strata(dat11cc) %>% select(Region, Year))
table(strata(wmncc) %>% select(Region, Year))
```

# Analysis of Molecular Variance

First, we should see what AMOVA tells us about population structure

```{r wmn_amova, cache = TRUE}
bdwmn <- bruvo.dist(wmncc, replen = other(wmncc)$REPLEN)
(wmn.amova <- poppr.amova(wmncc, ~Region/Year, dist = bdwmn))
(wmn.amova.test <- randtest(wmn.amova, nrepet = 999))
make_amova_table(wmn.amova, wmn.amova.test, samples = "Year")
```
```{r, full_amova, cache = TRUE}
bdfull <- bruvo.dist(dat11cc, replen = other(dat11cc)$REPLEN)
(full.amova.2 <- poppr.amova(dat11cc, ~Region/Year, dist = bdfull))
(full.amova.2.test <- randtest(full.amova.2, nrepet = 999))
(full.amova.3 <- poppr.amova(dat11cc, ~Region/SourceType/Year, dist = bdfull))
(full.amova.3.test <- randtest(full.amova.3, nrepet = 999))
make_amova_table(full.amova.2, full.amova.2.test, samples = "Year")
make_amova_table(full.amova.3, full.amova.3.test, samples = "Year")
```

```{r AMOVA_table, comment = NA}
make_amova_table(full.amova.3, full.amova.3.test, samples = "Year") %>%  
  as_tibble() %>%
  add_column(Hierarchy = c("Between Region", "Between Source within Region", "Between Year within Source", "Within Year"), .before = 1) %>%
  readr::write_csv(path = file.path(PROJHOME, "results", "tables", "AMOVA-year.csv"), col_names = TRUE) %>%
  rename(ps = `Phi statistic`) %>%
  mutate_if(is.numeric, format, digits = 3) %>%
  # mutate(ps = signif(ps, digits = 3)) %>%
  mutate(ps = gsub("0\\.00(\\d{1})(\\d{2})", "\\1.\\2e^-3^", ps)) %>%
  # mutate(ps = case_when(P > 0.05 ~ ps, TRUE ~ paste0("**", ps, "**"))) %>%
  rename(`$\\Phi$ statistic` = ps) %>%
  rename(`% variation` = `Percent variation`) %>%
  rename(S.S. = `Sum of Squares`) %>%
  rename(`*P*` = P) %>%
  select(c(1:4, 6, 5)) %>%
  huxtable::as_huxtable(add_colnames = TRUE) %>% 
  huxtable::set_col_width(c(0.8, 0.15, 0.15, 0.35, 0.5, 0.15)) %>% 
  huxtable::set_align(huxtable::everywhere, 2:6, "right") %>% 
  huxtable::set_number_format(huxtable::everywhere, 2, 0) %>%
  huxtable::set_number_format(huxtable::everywhere, 6, 3) %>%
  huxtable::set_number_format(huxtable::everywhere, 5, 4) %>%
  huxtable::print_md(max_width = 80)
```

# DAPC

Now that we have a result for the AMOVA showing a signficant difference in years, we can use DAPC to see where this difference is.

```{r dapc}
setPop(dat11cc) <- ~Region/Year
regyear.dapc <- dapc(dat11cc, n.pca = 20, n.da = 20)
```

```{r dapc_plot, fig.width = 7, fig.height = 7, warning = FALSE, message = FALSE}
# A ggplot2 version of scatter.dapc
#
# @param DAPC an object of class "dapc" derived from [adegenet::dapc]
# @param STRATA a data frame defining the population strata (see [adegenet::strata])
# @param color the variable in `STRATA` defining the color palette for the plot
# @param filter a "quosure" containing a filtering method to be passed to [dplyr::filter]
#
# @return
# @export
#
# @examples
# library("adegenet")
# data(microbov)
# strata(microbov) <- as.data.frame(other(microbov))
# setPop(microbov) <- ~spe/breed
# mscat <- ggscatter(dapc(microbov, n.pca = 20, n.da = 40), strata(microbov),
#                    color = "breed",
#                    filter = quo(breed %in% c("Borgou", "Zebu", "Montbeliard", "Salers")))
# mscat + facet_wrap(~coun)
# mscat + facet_wrap(~spe)
# mscat + facet_wrap(~breed)
ggscatter <- function(DAPC, STRATA, color = "Year", filter = NULL){
  RYD <- bind_cols(Population = DAPC$grp, STRATA, as.data.frame(DAPC$ind.coord)) %>%
    as_tibble()
  RYD_pop <- RYD %>% 
    group_by(Population) %>% 
    summarize_if(is.numeric, mean, na.rm = TRUE) %>%
    rename_all(function(x) gsub("LD", "mean", x))
  RYD <- full_join(RYD, RYD_pop, by = "Population")
  yminor <- pretty(RYD$LD2)
  xminor <- pretty(RYD$LD1)
  RYD <- if (!is.null(filter)) filter(RYD, !!filter) else RYD
  RYD_PLOT <- ggplot(RYD, aes_string(x = "LD1", y = "LD2", color = color)) + 
    geom_text(aes_string(label = color), alpha = 0.75) +
    geom_segment(aes(x = mean1, y = mean2, xend = LD1, yend = LD2), alpha = 0.5) +
    stat_ellipse(type = "norm", level = 0.66, alpha = 0.75) + 
    theme_bw(base_size = 16, base_family = "Helvetica") +  
    theme(aspect.ratio = 0.8) +
    theme(legend.position = "bottom") +
    theme(axis.text = element_blank()) + 
    theme(axis.title = element_blank()) + 
    theme(axis.ticks = element_blank()) + 
    viridis::scale_color_viridis(discrete = TRUE, option = "C", direction = -1) +
    viridis::scale_fill_viridis(discrete = TRUE,  option = "C", direction = -1) +
    scale_y_continuous(breaks = 0, minor_breaks = yminor) + 
    scale_x_continuous(breaks = 0, minor_breaks = xminor) + 
    theme(panel.background = element_rect(fill = "grey95")) +
    theme(panel.grid.major = element_line(color = "grey20")) +
    theme(panel.grid.minor = element_line(color = "white")) 
  RYD_PLOT
}

par(mfrow = c(2, 1))
loadingplot(regyear.dapc$var.contr, axis = 1)
loadingplot(regyear.dapc$var.contr, axis = 2)
par(mfrow = c(1, 1))

gg_region_year <- ggscatter(regyear.dapc, strata(dat11cc)) +
  facet_wrap(~Region) +
  theme(legend.position = c(0.75, 0.1)) +
  guides(color = guide_legend(nrow = 4)) +
  theme(legend.direction = "horizontal") +
  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
  theme(panel.border = element_blank())
gg_region_year
if (!interactive()) {
  ggsave(filename = file.path(PROJHOME, "results/figures/publication/dapc_region_year.pdf"),
         plot = gg_region_year,
         width = 7,
         height = 7)
}

```

The loading plot shows us the variables that are important for the first axis of
separation, mainly:

----------------
 Locus    Allele
-------- -------
7-2(F)       174
8-3(H)       256
9-2(F)       372
55-4(F)      157
110-4(H)     370
----------------

There's an important thing going on in Washington where it appears that the
2008 population is separated from the main population.

```{r dapc_plot_newamica, fig.width = 5, fig.height = 4, warning = FALSE, message = FALSE}
gg_region_year_micanewa <- ggscatter(regyear.dapc, strata(dat11cc), filter = quo(Region %in% c("WA", "CA", "NE", "MI"))) +
  facet_wrap(~Region, nrow = 2) +
  theme(legend.position = "right") +
  theme(legend.justification = "bottom") +
  theme(legend.box.margin = unit(c(0, 0, 0, 0), "lines")) +
  theme(legend.margin = unit(c(0, 0, 0, 0), "lines")) +
  theme(strip.background = element_rect(color = NA, fill = "grey90")) +
  theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
  theme(panel.border = element_blank()) +
  theme(legend.key = element_rect(fill = "grey95"))
gg_region_year_micanewa
if (!interactive()) {
    ggsave(filename = file.path(PROJHOME, "results/figures/publication/dapc_region_year_micanewa.pdf"),
       plot = gg_region_year_micanewa,
       width = 88,
       height = 0.8*88,
       units = "mm",
       scale = 1.25
       )
}
```

If we look at the tables above, we can see that there are only 12 samples from 
2008, all of whicha are in white mold screening nurseries. We also notice that
the coördinates for these isolates in the first two discriminant components are
overlapping with CA isolates. I suspect that there may be duplicate genotypes:

```{r crosspop, results = "asis"}
mlg.crosspop(dat11cc, ~Region/Year, quiet = TRUE, df = TRUE) %>%
  inner_join(filter(., Population == "WA_2008") %>% select(MLG), by = "MLG") %>%
  spread(Population, Count, fill = 0) %>%
  knitr::kable()
```

Well! It appears that ALL of the isolates in WA in 2008 have buddies in 2005 in
CA and France. We can take a look at how the allele frequencies are responding.

```{r fig.width = 10, fig.height = 12}
dat11ccloc <- dat11cc[loc = c("7-2(F)", "8-3(H)", "9-2(F)", "55-4(F)", "110-4(H)")]
loc <- map(seploc(dat11ccloc), . %>% 
             tab() %>% 
             as.data.frame() %>% 
             bind_cols(strata(dat11cc)) %>% 
             select(-MCG) %>%
             group_by(Region, Year) %>% 
             mutate(N = n()) %>%
             filter(N > 5) %>%
             summarize_if(is.numeric, mean, na.rm = TRUE) %>%
             ungroup() %>%
             mutate(Year = as.integer(as.character(Year))) %>%
             gather(allele, frequency, -Region, -Year, -N) %>%
             mutate(allele = as.integer(gsub("^.+?\\.([[:alnum:]]+)$", "\\1", allele)))
             # ,
             # .id = "Locus"
           )

ggloc <- function(dat, REG, years = 2003:2008){
  map(dat, . %>% 
    filter(Region == REG) %>% {
      ggplot(., aes(x = Year, y = frequency, color = allele, group = allele)) +
        geom_text(aes(label = paste0("(", N, ")")), y = 0.99, color = "grey20", vjust = 1) +
        geom_line(alpha = 0.75) +
        geom_point(aes(size = N), show.legend = FALSE, alpha = 0.75) +
        scale_x_continuous(breaks = years) +
        scale_y_continuous(limits = c(0, 1)) +
        theme_bw(base_size = 16, base_family = "Helvetica") +
        theme(legend.position = "top") +
        theme(aspect.ratio = 1/2) +
        theme(axis.text = element_text(color = "black")) +
        theme(axis.ticks.y = element_blank()) +
        theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
        # theme(axis.text.x.top = element_text(angle = 90, vjust = 0.5, hjust = 0)) +
        theme(panel.grid.minor.x = element_line(linetype = 0, color = "grey50")) +
        theme(panel.grid.major.x = element_line(linetype = 0, color = "grey50")) +
        theme(panel.grid.major = element_line(colour = "grey20")) +
        theme(panel.grid.minor = element_line(linetype = 3, colour = "grey50")) +
        # theme(panel.spacing.y = unit(0, "line")) +
        theme(panel.background = element_rect(color = NA, fill = "grey98")) +
        theme(strip.text = element_text(face = "bold", hjust = 0.05)) +
        theme(panel.border = element_blank()) +
        scale_size(range = c(3, 6), limits = c(6, 25)) +
        viridis::scale_color_viridis(option = "C", end = 0.9, 
                                     guide = "legend",
                                     breaks = sort(unique(.$allele)))
  })
}
ggl <- suppressMessages(ggloc(loc, "WA"))

cowplot::plot_grid(plotlist = ggl, labels = names(ggl), ncol = 2)
```

We can see that the allele frequencies seem to have a drastic change in the 
allele frequencies from 2007 to 2008. What happens if we look at a stable region
like NE?

```{r NEplot, fig.width = 10, fig.height = 10}
gglNE <- suppressMessages(ggloc(loc, "NE", 2003:2012))
cowplot::plot_grid(plotlist = gglNE, labels = names(gglNE), ncol = 2)
```
```{r MIplot, fig.width = 10, fig.height = 10}
gglMI <- suppressMessages(ggloc(loc, "MI", 2003:2012))
cowplot::plot_grid(plotlist = gglMI, labels = names(gglMI), ncol = 2)
```

It appears that MI is fairly stable, NE has some weird things going on with 2007
and 2008... I'm wondering if this may be a labeling issue.

<details>
<summary>Session Information</summary>

```{r, echo = FALSE}
options(width = 100)
devtools::session_info()
```

</details>