### 2017-10-02 README.Rmd
Create Rmd file for README

---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Analysis of 366 *S. sclerotiorum* isolates

[![Last-changedate](https://img.shields.io/badge/last%20change-`r gsub('-', '--', Sys.Date())`-brightgreen.svg)](https://github.com/everhartlab/sclerotinia-366/commits/master) [![minimal R version](https://img.shields.io/badge/R%3E%3D-3.4.1-brightgreen.svg)](https://cran.r-project.org/) [![Licence](https://img.shields.io/github/license/mashape/apistatus.svg)](http://choosealicense.com/licenses/mit/)  [![Circle CI](https://circleci.com/gh/everhartlab/sclerotinia-366.svg?style=shield&circle-token=:circle-token)](https://circleci.com/gh/everhartlab/sclerotinia-366)

This repository contains data, code, and a manuscript for analysis of 366 isolates of *Sclerotinia sclerotiorum* from the US and various countries around the world.

> Population structure and phenotypic variation of *Sclerotinia sclerotiorum*
> from dry bean in the United States
>
> Z. N. Kamvar, B. S. Amaradasa, R. Jhala, S. McCoy, J. Steadman, and S. E. Everhart

# TOC

The analyses are arranged in the following order according to the [Makefile]:

 1. [table-1.md]
 1. [MCG-virulence.md]
 1. [locus-stats.md]
 1. [MLG-distribution.md]
 1. [mlg-mcg.md]
 1. [RDA-analysis.md]
 1. [pop-diff.md]
 1. [tree.md]
 1. [wmn-differentiation.md]
 1. [by-year.md]

# Reproducing the analysis

## Locally

This project is controlled via a [Makefile] which means that everything (analyses, tables, figures, the paper itself) is controlled via one command:

```
make
```

This will bootstrap the installation (warning: it will update packages),
process the data, perform the analyses, and compile the paper.

Required software:

 - GNU Make (If you're on Windows, you can use MinGW: http://www.mingw.org/)
 - [R (version 3.4.1 or greater)](https://r-project.org)
 - [LaTeX](https://www.latex-project.org/get)
 - [pandoc](http://pandoc.org/) (Note: pandoc ships with Rstudio)
 - [devtools](https://github.com/hadley/devtools#readme)

## Docker

This repository contains a [Dockerfile](Dockerfile), which specifies the
instructions to build a [docker](https://www.docker.com/) container. This is
designed to capture the complete development environment of the analysis so
that it can be accurately reproduced. The image is ~2.71Gb, so be sure that you
have enough memory on your computer to run it.

To Install Docker, go here: https://docs.docker.com/engine/installation/#desktop. Once you have downloaded docker, you can either pull the container or build it. Pulling is by far the quickest way to do this.
The docker container is located at https://hub.docker.com/r/zkamvar/sclerotinia-366/

> Note: both ways assume that you are in the analysis directory

### Pulling the container

```
docker run --rm -it -v $(pwd):/analysis zkamvar/sclerotinia-366:latest bash
```

### Building the container

If you don't want to pull from docker hub, you can build the container.

```sh
docker build .
```

This will bootstrap the environment to build the container. Once it's finished
you should see:

```
Successfully built 8e1e4cd82e19
```

where `8e1e4cd82e19` will be replaced by your hash. Now that things are built,
you can run the analysis in the image with:

```
docker run -it -v $(pwd):/analysis 8e1e4cd82e19 bash
```

### Running the Analysis

Once you are in the container, you can run the analysis, which is mapped to
`analysis/`. The `make clean` command will wipe out all derivative files and
the `make` command will generate everything. Note that this took almost 2 hours
to run on my machine due to several bootstrapping processes.

```
cd analysis/
make clean
make
```

[Makefile]: Makefile
[table-1.md]: results/table-1.md
[MCG-virulence.md]: results/MCG-virulence.md
[locus-stats.md]: results/locus-stats.md
[MLG-distribution.md]: results/MLG-distribution.md
[mlg-mcg.md]: results/mlg-mcg.md
[RDA-analysis.md]: results/RDA-analysis.md
[pop-diff.md]: results/pop-diff.md
[tree.md]: results/tree.md
[wmn-differentiation.md]: results/wmn-differentiation.md
[by-year.md]: results/by-year.md

## Packages Used

```{r}
options(width = 100)
imports <- packageDescription("WorldSclerotinia")$Imports
imports <- strsplit(imports, "[^A-z]*,\n")[[1]]
for (i in imports) suppressPackageStartupMessages(library(i, character.only = TRUE))
devtools::session_info()
```